---
title: "Bookmarking"
author: "Brenda M. Moreno"
date: "02/04/2023"
output:
  html_document: 
    toc: yes
    highlight: kate
    theme: flatly
    number_sections: no
runtime: shiny
---

# CHAPTER 11:

In this chapter we'll learn how to do a bookmarking with shiny. We begin by loading shiny and other library:

```{r, eval = TRUE}
library(shiny)
```

# Basic Idea
  Let’s take a simple app that we want to make bookmarkable. This app draws Lissajous figures, which replicate the motion of a pendulum. This app can produce a variety of interesting patterns that you might want to share.
```{r, eval = TRUE, echo = TRUE}
ui <- fluidPage(
  sidebarLayout(
    sidebarPanel(
      sliderInput("omega", "omega", value = 1, min = -2, max = 2, step = 0.01),
      sliderInput("delta", "delta", value = 1, min = 0, max = 2, step = 0.01),
      sliderInput("damping", "damping", value = 1, min = 0.9, max = 1, step = 0.001),
      numericInput("length", "length", value = 100)
    ),
    mainPanel(
      plotOutput("fig")
    )
  )
)

server <- function(input, output, session) {
  t <- reactive(seq(0, input$length, length.out = input$length * 100))
  x <- reactive(sin(input$omega * t() + input$delta) * input$damping ^ t())
  y <- reactive(sin(t()) * input$damping ^ t())
  output$fig <- renderPlot({
    plot(x(), y(), axes = FALSE, xlab = "", ylab = "", type = "l", lwd = 2)
  }, res = 96)
}

shinyApp(ui, server)
```

  There are three things we need to do to make this app bookmarkable:\
  1. Add a **bookmarkButton()** to the UI. This generates a button that the user clicks to generate the bookmarkable URL;\
  2. Turn ui into a function. You need to do this because bookmarked apps have to replay the bookmarked values: effectively, Shiny modifies the default value for each input control. This means there’s no longer a single static UI but multiple possible UIs that depend on parameters in the URL (i.e., it has to be a function);\
3. Add enableBookmarking = "url" to the shinyApp() call.

```{r, eval = TRUE, echo = TRUE}
ui <- function(request) {
  fluidPage(
    sidebarLayout(
    sidebarPanel(
      sliderInput("omega", "omega", value = 1, min = -2, max = 2, step = 0.01),
      sliderInput("delta", "delta", value = 1, min = 0, max = 2, step = 0.01),
      sliderInput("damping", "damping", value = 1, min = 0.9, max = 1, step = 0.001),
      numericInput("length", "length", value = 100),
      bookmarkButton()
    ),
    mainPanel(
      plotOutput("fig")

      )
    )
  )
}

server <- function(input, output, session) {
  t <- reactive(seq(0, input$length, length.out = input$length * 100))
  x <- reactive(sin(input$omega * t() + input$delta) * input$damping ^ t())
  y <- reactive(sin(t()) * input$damping ^ t())
  output$fig <- renderPlot({
    plot(x(), y(), axes = FALSE, xlab = "", ylab = "", type = "l", lwd = 2)
  }, res = 96)
}

shinyApp(ui, server, enableBookmarking = "url")
```

      My bookmarkikng doesn't working!!!!!!!!!!!!!!!!!!!!
# Updating the URL
```{r, eval = TRUE, echo = TRUE}
ui <- function(request) {
  fluidPage(
    sidebarLayout(
    sidebarPanel(
      sliderInput("omega", "omega", value = 1, min = -2, max = 2, step = 0.01),
      sliderInput("delta", "delta", value = 1, min = 0, max = 2, step = 0.01),
      sliderInput("damping", "damping", value = 1, min = 0.9, max = 1, step = 0.001),
      numericInput("length", "length", value = 100),
      bookmarkButton()
    ),
    mainPanel(
      plotOutput("fig")

      )
    )
  )
}

# Automatically bookmark every time an input changes
observe({
  reactiveValuesToList(input)
  session$doBookmark()
})
# Update the query string
onBookmarked(updateQueryString)

server <- function(input, output, session) {
  t <- reactive(seq(0, input$length, length = input$length * 100))
  x <- reactive(sin(input$omega * t() + input$delta) * input$damping ^ t())
  y <- reactive(sin(t()) * input$damping ^ t())

  output$fig <- renderPlot({
    plot(x(), y(), axes = FALSE, xlab = "", ylab = "", type = "l", lwd = 2)
}, res = 96)

  observe({
    reactiveValuesToList(input)
    session$doBookmark()
  })
  onBookmarked(updateQueryString)
}

shinyApp(ui, server, enableBookmarking = "url")
```

      My bookmarkikng doesn't working!!!!!!!!!!!!!!!!!!!!
