---
title: "Graphics"
author: "Brenda M. Moreno"
date: "02/04/2023"
output:
  html_document: 
    toc: yes
    highlight: kate
    theme: flatly
    number_sections: yes
---


# CHAPTER 7:

  This chapter is intended to teach how to make __interactive graphics__.
  
# Introduction:

  If you haven’t already installed Shiny, install it now with:
```{r, eval = FALSE} 
install.packages("shiny")
```

  Then load in your current R session:
```{r, eval = TRUE} 
library(shiny)
library(ggplot2)
``` 

# Basics:
  A plot can respond to four different mouse1 events: click, dblclick (double-click), hover (when the mouse stays in the same place for a little while), and brush (a 105 rectangular selection tool). 
  
```{r, eval = TRUE, echo = TRUE} 
ui <- fluidPage(
  plotOutput("plot", click = "plot_click"),
  verbatimTextOutput("info")
)

server <- function(input, output) {
  output$plot <- renderPlot({
    plot(mtcars$wt, mtcars$mpg)
  }, res = 96)
  
  output$info <- renderPrint({
    req(input$plot_click)
    x <- round(input$plot_click$x, 2)
    y <- round(input$plot_click$y, 2)
    cat("[", x, ", ", y, "]", sep = "")
  })
}

shinyApp(ui, server)
```

# Clicking:
    First we will study the nearPoints() with plot.
```{r, eval = TRUE, echo = TRUE} 
ui <- fluidPage(
  plotOutput("plot", click = "plot_click"),
  tableOutput("data")
)

server <- function(input, output, session) {
  output$plot <- renderPlot({
    plot(mtcars$wt, mtcars$mpg)
  }, res = 96)

  output$data <- renderTable({
    nearPoints(mtcars, input$plot_click, xvar = "wt", yvar = "mpg")
  })
}

shinyApp(ui, server)
```

  And than we will study the nearPoints() with ggplot2.
```{r, eval = TRUE, echo = TRUE} 
ui <- fluidPage(
  plotOutput("plot", click = "plot_click"),
  tableOutput("data")
)

server <- function(input, output, session) {
  output$plot <- renderPlot({
    ggplot(mtcars, aes(wt, mpg)) + geom_point()
  }, res = 96)

  output$data <- renderTable({
    req(input$plot_click)
    nearPoints(mtcars, input$plot_click)
  })
}

shinyApp(ui, server)
```
  Another way to use nearPoints() is with allRows = TRUE and addDist = TRUE.
  That will return the original data frame with two new columns:
    • dist_ gives the distance between the row and the event (in pixels).
  • selected_ says whether or not it’s near the click event (i.e., whether or not it’s a row that would be returned when allRows = FALSE).

# Other Point Events:
    The same approach works equally well with click, dblclick, and hover: just change the name of the argument.

# Brushing:
  Another way of selecting points on a plot is to use a brush, a rectangular selection defined by four edges.
```{r, eval = TRUE, echo = TRUE} 
ui <- fluidPage(
  plotOutput("plot", brush = "plot_brush"),
  tableOutput("data")
)

server <- function(input, output, session) {
  output$plot <- renderPlot({
    ggplot(mtcars, aes(wt, mpg)) + geom_point()
  }, res = 96)

  output$data <- renderTable({
    brushedPoints(mtcars, input$plot_brush)
  })
}

shinyApp(ui, server)
```
  OBS: Use brushOpts() to control the color (fill and stroke), or restrict brushing to a single dimension with direction = "x" or "y".

# Modifying the Plot:
  As you might guess from the name, reactiveVal() is rather similar to reactive(). You create a reactive value by calling reactiveVal() with its initial value, and retrieve that value in the same way as a reactive:
```{r, eval = FALSE, echo = TRUE} 
val <- reactiveVal(10)
val()
```

  The big difference is that you can also update a reactive value, and all reactive consumers that refer to it will recompute. A reactive value uses a special syntax for updating—you call it like a function, with the first argument being the new value:
```{r, eval = FALSE, echo = TRUE} 
val(20)
val()
```
  Unfortunately, if you actually try to run this code in the console, you’ll get an error because it has to be run in a reactive environment. That makes experimentation an debugging more challenging because you’ll need to use browser() or similar to pause execution within the call to shinyApp().
```{r, eval = TRUE, echo = TRUE} 
set.seed(1014)
df <- data.frame(x = rnorm(100), y = rnorm(100)) #??

ui <- fluidPage(
  plotOutput("plot", click = "plot_click", )
)

server <- function(input, output, session) {
  dist <- reactiveVal(rep(1, nrow(df)))
  observeEvent(input$plot_click,
    dist(nearPoints(df, input$plot_click, allRows = TRUE, addDist = TRUE)$dist_)
)
  
  output$plot <- renderPlot({
    df$dist <- dist()
    ggplot(df, aes(x, y, size = dist)) +
      geom_point() +
      scale_size_area(limits = c(0, 1000), max_size = 10, guide = NULL)
    }, res = 96)
}

shinyApp(ui, server)
# This app uses a reactiveVal() to store the distance to the point that was last clicked, which is then mapped to point size. Here I show the results of clicking on a point on the far left.
```

  Here’s a more complicated idea. I want to use a brush to progressively add points to a selection. Here I display the selection using different colors, but you could imagine many other applications. 
```{r, eval = TRUE, echo = TRUE} 
ui <- fluidPage(
  plotOutput("plot", brush = "plot_brush", dblclick = "plot_reset")
)

server <- function(input, output, session) {
  selected <- reactiveVal(rep(FALSE, nrow(mtcars)))

  observeEvent(input$plot_brush, { # ?brush?
    brushed <- brushedPoints(mtcars, input$plot_brush, allRows = TRUE)$selected_
    selected(brushed | selected())
  })
  observeEvent(input$plot_reset, {
    selected(rep(FALSE, nrow(mtcars)))
  })

  output$plot <- renderPlot({
    mtcars$sel <- selected()
    ggplot(mtcars, aes(wt, mpg)) +
      geom_point(aes(colour = sel)) +
      scale_colour_discrete(limits = c("TRUE", "FALSE"))
    }, res = 96)
}

shinyApp(ui, server)

```

# Dynamic Height and Width:

```{r, eval = TRUE, echo = TRUE} 
ui <- fluidPage(
  sliderInput("height", "height", min = 100, max = 500, value = 250),
  sliderInput("width", "width", min = 100, max = 500, value = 250),
  plotOutput("plot", width = 250, height = 250)
)

server <- function(input, output, session) {
  output$plot <- renderPlot(
    width = function() input$width,
    height = function() input$height,
    res = 96,
    {
      plot(rnorm(20), rnorm(20))
    }
  )
}

shinyApp(ui, server)

```

# Images:
  You can use renderImage() if you want to display existing images (not plots). For example, you might have a directory of photographs that you want shown to the user -> For add the logo of app.
```{r, eval = TRUE, echo = TRUE} 
puppies <- tibble::tribble(
  ~breed, ~ id, ~author,
  "corgi", "eoqnr8ikwFE","alvannee",
  "labrador", "KCdYn0xu2fU", "shaneguymon",
  "spaniel", "TzjMd7i5WQI", "_redo_"
)

ui <- fluidPage(
  selectInput("id", "Pick a breed", choices = setNames(puppies$id, puppies$breed)),
  htmlOutput("source"),
  imageOutput("photo")
)

server <- function(input, output, session) {
  output$photo <- renderImage({
    list(
      src = file.path("puppy-photos", paste0(input$id, ".jpg")),
      contentType = "image/jpeg",
      width = 500,
      height = 650
    )
  }, deleteFile = FALSE)

  output$source <- renderUI({
    info <- puppies[puppies$id == input$id, , drop = FALSE]
    HTML(glue::glue("<p>
<a href='https://unsplash.com/photos/{info$id}'>original</a> by
<a href='https://unsplash.com/@{info$author}'>{info$author}</a>
</p>"))
  })
}

shinyApp(ui, server)

```  
